Pleiades: Interactive Composing Tools for Vega-Lite Charts

### 2019-06-12 final/paper/paper.tex
+Result +Discussion

In practice, the above approach is proven to be too rigid. One example of
unintended consequences of the above approach is that users can’t edit any
specification at all, because, during the editing, it’s very likely that
the text won’t be able to parse correctly into the JSON format. We thus
separate operations into macro-operations and micro-operations, where a
macro-operation is a group of micro-operations. We then relax our goal to
only apply to macro-operations. That is, every macro-operation preserves
well-formedness. For micro-operations that does not preserve well-formedness,
the operation won’t affect the working area directly. Rather, it is confined
inside a dialog box so that a batch of micro-operations can be readily reversed.

@@

\paragraph{The benefit of using syntax tree} With the syntax tree, inner views
can be located to modify easily. Furthermore, when the working view is output to
the working area, it can be rendered recursively through each node in the tree.
This recursive rendering allows us to render inner views as separate views from
the outer view. For example, when rendering concat view, the inner views are
rendered separately in a way that users can select either the inner view or
the whole concat view.

\subsection{How data of each spec is pulled to the top level in facet and repeat}
In facet and repeat spec, data must be at the topmost level of the output JSON
spec. When RepeatView or FacetView is exported to JSON, Pleiades pull out all
the data source in every children view of the inner view to the top level. If
there is more than one distinct data source, Pleiades will use the first one it
encounters by pre-order tree traversal. Pleiades also uses a similar method to
exhaustively search for every data source in the inner views of an operand to
validate if there should be a warning shown for the current operand for facet/repeat.

@@

result here

@@

discussion here

>>>>

In practice, the above approach is proven to be too rigid. One example of
unintended consequences of the above approach is that users can’t edit any
specification at all, because, during the editing, it’s very likely that
the text won’t be able to parse correctly into the JSON format. We thus
separate operations into macro-operations and micro-operations, where a
macro-operation is a group of micro-operations. We then relax our goal to
only apply to macro-operations. That is, every macro-operation preserves
well-formedness. For micro-operations that does not preserve well-formedness,
the operation won’t affect the working area directly. Rather, it is confined
inside a dialog box so that a batch of micro-operations can be readily reverted.

@@

\paragraph{The benefit of syntax tree} With the concept of syntax tree, inner views
can be located to modify easily. When the working view is output to
the working area, it can be rendered recursively through each node in the tree.
This recursive rendering allows us to render inner views as separate views from
the outer view. For example, when rendering concat view, the inner views are
rendered separately in a way that users can select either the inner view or
the whole concat view.

\subsection{Data Pulling}
In facet and repeat spec, data must be at the topmost level of the output JSON
spec. When RepeatView or FacetView is exported to JSON, Pleiades pull out all
the data source in every children view of the inner view to the top level. If
there is more than one distinct data source, Pleiades will use the first one it
encounters by pre-order tree traversal. Pleiades also uses a similar method to
exhaustively search for every data source in the inner views of an operand to
validate if there should be a warning shown for the current operand for facet/repeat.

@@

To evaluate our system, we perform a case study aiming to answer the following
questions:

\begin{itemize}
\item How easy is it to use Pleiades?
\item Does Pleiades allow effective view composition?
\end{itemize}

For the question how easy it is to use Pleiades, two participants of the case
study report back that the tool is mostly easy to use. They wish that
the tool provides more visual cues and that they prefer drag and drop to
operand/operator selection, however. Some of these suggestions are already
incorporated into the final version of the tool.

For the question whether Pleiades allows effective view composition,
the participants' feedback is highly positive, stating that charts
can be composited in an intuitive way and visually responsive.

During our poster presentation, we additionally asked a lot of our peers
to try out our application and compare their workflow to that when they used the
Vega-Lite API along with a JSON spec. The feedback is similar to what we have
seen in the case study.

Based on these results, we conclude that Pleiades eases beginners to
develop composite charts, save time for users, ease data-exploration,
and act as a support toolkit for education.

@@

As we expected, the results show that Pleiades gives users a better understanding
and ease them at composing different charts and see what fits the best.
The key idea behind why Pleiades makes composing charts
easier and more intuitive is because the process allows users to compose small
charts and rearrange them into a bigger charts. We call this approach
\emph{bottom-up}. This contrasts with Vega-lite's \emph{top-down approach}
which demands users to know the layout of the entire structure already
beforehand and hence is very rigid and inflexible. Also, the graphical user interface
provided used with a smoother way to compose charts, than to write nested JSON specs.

One interesting result we found was that the ability to perform
the four operations gave users a better understanding of these operations
and how to properly use then. The validation steps, along with the error message
that Pleiades provides helps users to understand what they are doing wrong,
helping them in their education about visualization in general.

In terms of exploring data, the time it takes to compose different
charts and see find useful results, went down by a lot because of the
high-level of interactivity provided. Users don’t have to worry about
updating JSON code for every small change in their chart. Moreover,
a bad layout could be a hindrance that shows interesting trends in a chart.
Our tool which allows users to experiment with composition easily thus helps
with data exploration too.
